decompr: GVC decomposition in R

### 2015-01-22 draft.Rnw
Update draft.Rnw
Various edits pt. II

\vspace{\stretch{1}}
\abstract{Global Value Chains have become a central unit of analysis in research on international trade. However, 
the complex matrix transformations at the basis of most Value Chain indicators still constitute a significant entry barrier
the field. The R package \textit{decompr} solves this problem by implementing algorithms for the analysis of Global Value Chains,
which calculate standard indicators in an automated fashion. Two methods for gross export flow decomposition using 
Inter-Country Input-Output tables are provided. The first method is the \textit{Wang-Wei-Zhu} algorithm which splits 
bilateral gross exports into 16 value added components. These components can broadly be divided into domestic and 
foreign value added. The second method concerns a decomposition based on the classical \citet{wale36} insight.
It derives the value added origins of an industry's exports by source country and source industry, using easily available
gross trade data. This article summarises the methodology of the algorithms, describes the format of the input and output data, 
and exemplifies the usefulness of the two methods on the basis of a simple example data set.}
\pagebreak

@@

\begin{equation}
VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...),
\end{equation}
which as an infinite geometric series with the elements of $A<1$ simplifies to

\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textit{V} is a N x N matrix with the diagonal representing the direct value added contribution of the industries, \textit{A} is the Input-Output coefficient matrix with dimension N x N, i.e. it gives the direct input flows between industries required for 1\$ of output, $B = (I-A)^{-1}$ is then the so called Leontief inverse, and \textit{T} indicates a matrix transpose operation. \textit{VB} gives thus a N x N matrix of so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries. Looking from the perspective of the supplying industries, the matrix gives the value added origins of an industry's production. If we multiply it with a N x N matrix whose diagonal specifies each industry's total output, we get value added origins as absolute values instead of shares.

@@

\begin{align}
\begin{split}
V (I- & A)^{-1} E =
\begin{pmatrix}
v_{i}^{k}& 0& 0& 0\\
0& v_{j}^{k}& 0& 0\\
0& 0& v_{i}^{l}& 0\\
0& 0& 0& v_{j}^{l}
\end{pmatrix}
*
\begin{pmatrix}
b_{ii}^{kk}& b_{ij}^{kk}& b_{ii}^{kl}& b_{ij}^{kl}\\
b_{ji}^{kk}& b_{jj}^{kk}& b_{ji}^{kl}& b_{jj}^{kl}\\
b_{ii}^{lk}& b_{ij}^{lk}& b_{ii}^{ll}& b_{ij}^{ll}\\
b_{ji}^{lk}& b_{jj}^{lk}& b_{ji}^{ll}& b_{jj}^{ll}
\end{pmatrix}
*
\begin{pmatrix}
e_{i}^{k}& 0& 0& 0\\
0& e_{j}^{k}& 0& 0\\
0& 0& e_{i}^{l}& 0\\
0& 0& 0& e_{j}^{l}
\end{pmatrix}
=
\\
&
\begin{pmatrix}
v_{i}^{k}b_{ii}^{kk}e_{i}^{k}& v_{i}^{k} b_{ij}^{kk}e_{j}^{k}& v_{i}^{k}b_{ii}^{kl}e_{i}^{l}& v_{i}^{k}b_{ij}^{kl}e_{j}^{l}\\
v_{j}^{k}b_{ji}^{kk}e_{i}^{k}& v_{j}^{k}b_{jj}^{kk}e_{j}^{k}& v_{j}^{k}b_{ji}^{kl}e_{i}^{l}& v_{j}^{k}b_{jj}^{kl}e_{j}^{l}\\
v_{i}^{l}b_{ii}^{lk}e_{i}^{k}& v_{i}^{l}b_{ij}^{lk}e_{j}^{k}& v_{i}^{l}b_{ii}^{ll}e_{i}^{l}& v_{i}^{l}b_{ij}^{ll}e_{j}^{l}\\
v_{j}^{l}b_{ji}^{lk}e_{i}^{k}& v_{j}^{l}b_{jj}^{lk}e_{j}^{k}& v_{j}^{l}b_{ji}^{ll}e_{i}^{l}& v_{j}^{l}b_{jj}^{ll}e_{j}^{l}
\end{pmatrix}
=
\begin{pmatrix}
vax_{ii}^{kk}& vax_{ij}^{kk}& vax_{ii}^{kl}& vax_{ij}^{kl}\\
vax_{ji}^{kk}& vax_{jj}^{kk}& vax_{ji}^{kl}& vax_{jj}^{kl}\\
vax_{ii}^{lk}& vax_{ij}^{lk}& vax_{ii}^{ll}& vax_{ij}^{ll}\\
vax_{ji}^{lk}& vax_{jj}^{lk}& vax_{ji}^{ll}& vax_{jj}^{ll}
\end{pmatrix}
\end{split}
\end{align}
with
\begin{align*}
&v_{s}^{c} = \frac{va_{s}^{c}}{y_{s}^{c}} = 1 - \sum_{d}^{2}{a_{kc}^{ds}} - \sum_{d}^{2}{a_{lc}^{ds}} \hspace{2cm} (c \in k,l \hspace{.5cm} s \in i,k),\\
&\begin{pmatrix}
b_{ii}^{kk}& b_{ij}^{kk}& b_{ii}^{kl}& b_{ij}^{kl}\\
b_{ji}^{kk}& b_{jj}^{kk}& b_{ji}^{kl}& b_{jj}^{kl}\\
b_{ii}^{lk}& b_{ij}^{lk}& b_{ii}^{ll}& b_{ij}^{ll}\\
b_{ji}^{lk}& b_{jj}^{lk}& b_{ji}^{ll}& b_{jj}^{ll}
\end{pmatrix}
=
\begin{pmatrix}
1-a_{ii}^{kk}& -a_{ij}^{kk}& -a_{ii}^{kl}& -a_{ij}^{kl}\\
-a_{ji}^{kk}& 1-a_{jj}^{kk}& -a_{ji}^{kl}& -a_{jj}^{kl}\\
-a_{ii}^{lk}& -a_{ij}^{lk}& 1-a_{ii}^{ll}& -a_{ij}^{ll}\\
-a_{ji}^{lk}& -a_{jj}^{lk}& -a_{ji}^{ll}& 1-a_{jj}^{ll}
\end{pmatrix}^{-1},
\end{align*}
and
\begin{equation*}
a_{cf}^{su} = \frac{inp_{cf}^{su}}{y_{f}^{u}}  \hspace{2cm} (c,f \in k,l \hspace{.5cm} s,u \in i,k),
\end{equation*}
where $v_{s}^{c}$ gives the share of value added, $va_{s}^{c}$, in output , $y_{s}^{c}$, and $e_{s}^{c}$ indicates gross exports. Finally, $a_{cf}^{su}$ denotes the share of inputs , $inp_{cf}^{su}$, in output. The elements of the $V(I-A)^{-1}E$ or $vax$ matrix are our estimates for the country-industry level value added origins of each country-industry's exports. decompr implements this algorithm into R to automate the process of deriving the matrix. Equipped with it, researchers can calculate standard GVC indicators. Examples include \citet{dahuetal01}'s Vertical Specialisation ratio at the industry-level which can be derived by summing for each country and industry across the value added of all foreign countries and industries. Alternatively, the four dimensions of the matrix (source country, source industry, using country, using industry) allow for industry-level gravity-type estimations of value added trade flows.

@@

A key advantage of the decomposition becomes clear when we compare the decomposed values with the intermediate trade values of the non-decomposed IO table when multiplied with the exports over output ratio to create comparability (Table \ref{tab:noleon}). We see for instance that Argentina's Agriculture industry contributes significantly more value added to the German Transport Equipment industry than suggested by the non-decomposed IO table. The reason is that Argentina Agriculture industry is an important supplier to Turkey's Textile and Leather industry which is in turn an important supplier for the German Transport Equipment industry. The decomposition thus allows us to see how the value added flows along this Global Value Chain.

>>>>

\vspace{\stretch{1}}
\abstract{Global Value Chains have become a central unit of analysis in research on international trade. However, 
the complex matrix transformations at the basis of most Value Chain indicators still constitute a significant entry barrier
the field. The R package \textit{decompr} solves this problem by implementing algorithms for the analysis of Global Value Chains,
which calculate standard indicators in an automated fashion. Two methods for gross export flow decomposition using 
Inter-Country Input-Output tables are provided. The first method is the \textit{Wang-Wei-Zhu} algorithm which splits 
bilateral gross exports into 16 value added components. These components can broadly be divided into domestic and 
foreign value added in exports. The second method concerns a decomposition based on the classical \citet{wale36} insight.
It derives the value added origins of an industry's exports by source country and source industry, using easily available
gross trade data. This article summarises the methodology of the algorithms, describes the format of the input and output data, 
and exemplifies the usefulness of the two methods on the basis of a simple example data set.}
\pagebreak

@@

\begin{equation}
VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...),
\end{equation}
which, as an infinite geometric series with the elements of $A<1$, simplifies to

\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textit{V} is a N x N matrix with the diagonal representing the direct value added contribution of the industries,
\textit{A} is the Input-Output coefficient matrix with dimension N x N, i.e. it gives the direct input flows between
industries required for 1\$ of output, $B = (I-A)^{-1}$ is then the so called Leontief inverse, and \textit{T} indicates
a matrix transpose operation. \textit{VB} gives thus a N x N matrix of so called value added multipliers,
which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about
in all other industries. Looking from the perspective of the supplying industries, the matrix gives the value added that
they contribute to the using industry's production. If we multiply it with a N x N matrix whose diagonal specifies each
industry's total output, we get value added origins as absolute values instead of shares.

@@

\begin{align}
\begin{split}
V (I- A &)^{-1} E=
\begin{pmatrix}
v_{k}^{i}& 0& 0& 0\\
0& v_{k}^{j}& 0& 0\\
0& 0& v_{l}^{i}& 0\\
0& 0& 0& v_{l}^{j}
\end{pmatrix}
*
\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
*
\begin{pmatrix}
e_{k}^{i}& 0& 0& 0\\
0& e_{k}^{j}& 0& 0\\
0& 0& e_{l}^{i}& 0\\
0& 0& 0& e_{l}^{j}
\end{pmatrix}
=
\\
&
\begin{pmatrix}
v_{k}^{i}b_{kk}^{ii}e_{k}^{i}& v_{k}^{i} b_{kk}^{ij}e_{k}^{j}& v_{k}^{i}b_{kl}^{ii}e_{l}^{i}& v_{k}^{i}b_{kl}^{ij}e_{l}^{j}\\
v_{k}^{j}b_{kk}^{ji}e_{k}^{i}& v_{k}^{j}b_{kk}^{jj}e_{k}^{j}& v_{k}^{j}b_{kl}^{ji}e_{l}^{i}& v_{k}^{j}b_{kl}^{jj}e_{l}^{j}\\
v_{l}^{i}b_{lk}^{ii}e_{k}^{i}& v_{l}^{i}b_{lk}^{ij}e_{k}^{j}& v_{l}^{i}b_{ll}^{ii}e_{l}^{i}& v_{l}^{i}b_{ll}^{ij}e_{l}^{j}\\
v_{l}^{j}b_{lk}^{ji}e_{k}^{i}& v_{l}^{j}b_{lk}^{jj}e_{k}^{j}& v_{l}^{j}b_{ll}^{ji}e_{l}^{i}& v_{l}^{j}b_{ll}^{jj}e_{l}^{j}
\end{pmatrix}
=
\begin{pmatrix}
vae_{kk}^{ii}& vae_{kk}^{ij}& vae_{kl}^{ii}& vae_{kl}^{ij}\\
vae_{kk}^{ji}& vae_{kk}^{jj}& vae_{kl}^{ji}& vae_{kl}^{jj}\\
vae_{lk}^{ii}& vae_{lk}^{ij}& vae_{ll}^{ii}& vae_{ll}^{ij}\\
vae_{lk}^{ji}& vae_{lk}^{jj}& vae_{ll}^{ji}& vae_{ll}^{jj}
\end{pmatrix}
\end{split}
\end{align}
\begin{align*}
&v_{c}^{s} = \frac{va_{c}^{s}}{y_{c}^{s}} = 1 - a_{kc}^{is} - a_{kc}^{js} - a_{lc}^{js} - a_{lc}^{is} \hspace{2cm} (c \in k,l \hspace{.5cm} s \in i,j),\\
&\begin{pmatrix}
b_{kk}^{ii}& b_{kk}^{ij}& b_{kl}^{ii}& b_{kl}^{ij}\\
b_{kk}^{ji}& b_{kk}^{jj}& b_{kl}^{ji}& b_{kl}^{jj}\\
b_{lk}^{ii}& b_{lk}^{ij}& b_{ll}^{ii}& b_{ll}^{ij}\\
b_{lk}^{ji}& b_{lk}^{jj}& b_{ll}^{ji}& b_{ll}^{jj}
\end{pmatrix}
=
\begin{pmatrix}
1-a_{kk}^{ii}& -a_{kk}^{ij}& -a_{kl}^{ii}& -a_{kl}^{ij}\\
-a_{kk}^{ji}& 1-a_{kk}^{jj}& -a_{kl}^{ji}& -a_{kl}^{jj}\\
-a_{lk}^{ii}& -a_{lk}^{ij}& 1-a_{ll}^{ii}& -a_{ll}^{ij}\\
-a_{lk}^{ji}& -a_{lk}^{jj}& -a_{ll}^{ji}& 1-a_{ll}^{jj}
\end{pmatrix}^{-1},
\end{align*}
and
\begin{equation*}
a_{cf}^{su} = \frac{inp_{cf}^{su}}{y_{f}^{u}}  \hspace{2cm} (c,f \in k,l \hspace{.5cm} s,u \in i,j),
\end{equation*}
where \(v_{s}^{c}\) gives the share of industry \emph{s}'s value added,
\(va_{c}^{s},\) in output, \(y_{s}^{c}\), and \(e_{k}^{i}\)
indicates gross exports. Finally, \(a_{su}^{cf}\) denotes
the share of inputs, \(inp_{su}^{cf}\), in output. 
The elements of the $V(I-A)^{-1}E$ or $vae$ matrix are our estimates for the country-industry level value added
origins of each country-industry's exports. \textit{decompr} implements this algorithm into R to automate the process of
deriving the matrix. Equipped with it, researchers can calculate standard GVC indicators. 
Examples include \citet{dahuetal01}'s Vertical Specialisation ratio at the industry-level which can be derived
by summing for each country and industry across the value added of all foreign countries and industries.
Alternatively, the four dimensions of the matrix (source country, source industry, using country, using industry)
allow for industry-level gravity-type estimations of value added trade flows.

@@

A key advantage of the decomposition becomes clear when we compare the decomposed values with the intermediate trade
values of the non-decomposed IO table when multiplied with the exports over output ratio to create
comparability (Table \ref{tab:noleon}). We see for instance that Argentina's Agriculture industry
contributes significantly more value added to the German Transport Equipment industry than suggested
by the non-decomposed IO table. The reason is that Argentina's Agriculture industry is an important supplier
to Turkey's Textile and Leather industry which is in turn an important supplier for the German Transport
Equipment industry. The decomposition thus allows us to see how the value added flows along this Global Value Chain.
